\documentclass{book}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{hyperref}

\usepackage{fancyhdr,url}

%---------- fonts Type 1 -----------------
%\usepackage{times}
%\usepackage[T1]{fontenc}
\usepackage{textcomp}

%

%------------------------Page set-up-----------------------------------------

\renewcommand{\baselinestretch}{1.25}
\setlength{\hoffset}{-1in}
\setlength{\oddsidemargin}{3.5cm}
\setlength{\evensidemargin}{3.5cm}
\setlength{\topmargin}{0cm}
\setlength{\footskip}{2cm}
\setlength{\headheight}{14pt}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0cm}
\setlength{\parskip}{0.75\baselineskip}

%------------------------------------------------------------------------------


%---- change link style ----
\hypersetup{colorlinks, linkcolor=blue, pdfstartview={FitH}}


% Pages and Fancyheadings stuff
%-----------------------------------------------------------------------
\cfoot{\thepage}
\fancyhead[LE,RO]{}
\fancyhead[LO]{\nouppercase{\scshape\rightmark}}
\fancyhead[RE]{\nouppercase{\scshape\leftmark}}
%-----------------------------------------------------------------------

% boxes

\newsavebox{\fmboxb}
\newenvironment{mybox}
    {\vspace{-2mm}\begin{center}\begin{lrbox}{\fmboxb}\hspace{2mm}
        \begin{minipage}{0.85\textwidth} \vspace{2mm}\small}
    {  \vspace{2mm} \end{minipage}
    \hspace{2mm}\end{lrbox}\fbox{\usebox{\fmboxb}}\end{center}}



%----------------- TITLE --------------

\title{\Huge \bfseries SQLMap PHP DataMapper v1.0
    \thanks{Copyright 2006. All Rights Reserved.}
}
\author{Wei Zhuo}
\date{\today}


%-------------- BEGIN DOCUMENT ------------------


\begin{document}

\maketitle

\pagestyle{plain}
\addcontentsline{toc}{chapter}{Contents}
\pagenumbering{roman}
\tableofcontents

\chapter*{Legal Notice}
\addcontentsline{toc}{chapter}{Legal Notice}

Copies of this document may be made for your own use and for distribution to
others, provided that you do not charge any fee for such copies and further
provided that each copy contains this Copyright Notice, whether distributed in
print or electronically.

This document is largely based on the iBATIS.NET -- DataMapper Application
Framework Developer Guide.

\chapter*{License}
\addcontentsline{toc}{chapter}{License}
SQLMap for PHP is free software released under the terms of the following BSD license.\\
Copyright 2004-2006, PradoSoft (http://www.pradosoft.com)\\
All rights reserved.

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions are met:
\begin{enumerate}
    \item Redistributions of source code must retain the above copyright notice,
this list of conditions and the following disclaimer.

 \item Redistributions in binary form must reproduce the above copyright notice,
this list of conditions and the following disclaimer in the documentation
and/or other materials provided with the distribution.

\item Neither the name of the developer nor the names of its contributors may
be used to endorse or promote products derived from this software without
specific prior written permission.
\end{enumerate}

\begin{verbatim}
THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE
FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT
OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
\end{verbatim}


\newpage

\pagestyle{fancyplain}
\pagenumbering{arabic}

\include{ch1}
\include{ch2}
\include{ch3}
\include{ch4}
\include{ch5}
\include{ch6}
\include{ch7}
\include{ch8}
\include{ch9}

\end{document}
