\chapter{Using SQLMap PHP DataMapper}
The SQLMap DataMapper API provides four core functions:

\begin{itemize}
  \item build a \tt{TSqlMapper} instance from a configuration file or cache
  \item execute an update query (including insert and delete).
  \item execute a select query for a single object
  \item execute a select query for a list of objects
\end{itemize}

The API also provides support for retrieving paginated lists and managing
transactions.

\section{Building a \tt{TSqlMapper} instance}
An XML document is a wonderful tool for describing a database configuration
(Chapter~\ref{section:4.3}) or defining a set of data mappings
(Chapter~\ref{section:3}), but you can't execute XML. In order to use the
SQLMap configuration and definitions in your PHP application, you need a class
you can call.

The framework provides service methods that you can call which read the
configuration file (and any of its definition files) and builds a
\tt{TSqlMapper} object. The \tt{TSqlMapper} object provides access to the rest
of the framework. Example~\ref{example:9.4} shows a singleton Mapper that is
similar to the one bundled with the framework.

\begin{example}\label{example:9.4}
A Mapper singleton you can call from your own applications
\begin{verbatim}
require_once('/path/to/SQLMap/TSqlMapper.php');

class TMapper
{
    private static $_mapper;

    public static function configure($configFile)
    {
        if(is_null(self::$_mapper))
        {
            $builder = new TDomSqlMapBuilder();
            self::$_mapper = $builder->configure($configFile);
        }
        return self::$_mapper;
    }

    public static function instance()
    {
        return self::$_mapper;
    }
}
\end{verbatim}
\end{example}

To obtain the \tt{TSqlMapper} instance, first configure the mapper once,
\begin{verbatim}
TMapper::configure('path/to/sqlmap.config');
\end{verbatim}
The \tt{TDomSqlMapBuilder} object will go throught the the \tt{sqlmap.config}
file and build a \tt{TSqlMapper} instance. To use \tt{TSqlMapper} in your
application, specify one of the \tt{TSqlMapper} methods (see Section ???).
Here's an example:
\begin{verbatim}
$list = TMapper::instance()->queryForList("PermitNoForYearList", $values);
\end{verbatim}

\subsection{Multiple Databases}
If you need access to more than one database from the same application, create
a DataMapper configuration file for that database and another Mapper class to
go with it.

\subsection{\tt{TDomSqlMapBuilder} Configuration Options}
If you find that you already have loaded your DataMapper configuration
information as a \tt{SimpleXMLElement} instance within your application, the
\tt{TDomSqlMapBuilder} provides \tt{Configure} overloads for those types as
well.

\section{Exploring the SQLMap PHP DataMapper API through the \tt{TSqlMapper}}
The \tt{TSqlMapper} instance acts as a facade to provide access the rest of
the DataMapper framework. The DataMapper API methods are shown in Example
4.11.

\begin{example}
The SQLMap DataMapper API for PHP.
\begin{verbatim}
 /* Query API */
 public function queryForObject($statementName, $parameter=null, $result=null);
 public function queryForList($statementName, $parameter=null, $result=null,
                                    $skip=-1, $max=-1);
 public function queryForPagedList($statementName, $parameter=null, $pageSize=10);
 public function queryForMap($statementName, $parameter=null,
                                    $keyProperty=null, $valueProperty=null);

 public function insert($statementName, $parameter=null)
 public function update($statementName, $parameter=null)
 public function delete($statementName, $parameter=null)

 /* Connection API */
 public function openConnection()
 public function closeConnection()

 /* Transaction API */
 public function beginTransaction()
 public function commitTransaction()
 public function rollBackTransaction()
\end{verbatim}
\end{example}

Note that each of the API methods accept the name of the Mapped Statement as
the first parameter. The \tt{statementName} parameter corresponds to the
\tt{id} of the Mapped Statement in the Data Map definition (see
Section~\ref{section:3.3}). In each case, a \tt{parameterObject} also may be
passed. The following sections describe how the API methods work.

\subsection{Insert, Update, Delete}
\begin{verbatim}
 public function insert($statementName, $parameter=null)
 public function update($statementName, $parameter=null)
 public function delete($statementName, $parameter=null)
\end{verbatim}

If a Mapped Statement uses one of the \tt{<insert>}, \tt{<update>}, or
\tt{<delete>} statement-types, then it should use the corresponding API
method. The \tt{<insert>} element supports a nested \tt{<selectKey>} element
for generating primary keys (see Section~\ref{section:3.3.3}). If the
\tt{<selectKey>} stanza is used, then \tt{insert} returns the generated key;
otherwise a null object is returned. Both the \tt{update} and \tt{delete}
methods return the number of rows affected by the statement.

\subsection{QueryForObject}
\begin{verbatim}
public function queryForObject($statementName, $parameter=null, $result=null);
\end{verbatim}

If a Mapped Statement is expected to select a single row, then call it using
\tt{queryForObject}. Since the Mapped Statement definition specifies the
result class expected, the framework can both create and populate the result
class for you. Alternatively, if you need to manage the result object
yourself, say because it is being populated by more than one statement, you
can use the alternate form and pass your \tt{\$resultObject} as the third
parameter.

\subsection{QueryForList}

\begin{verbatim}
public function queryForList($statementName, $parameter=null, $result=null,
                                    $skip=-1, $max=-1);
\end{verbatim}
If a Mapped Statement is expected to select multiple rows, then call it using
\tt{queryForList}. Each entry in the list will be an result object populated
from the corresponding row of the query result. If you need to manage the
\tt{\$resultObject} yourself, then it can be passed as the third parameter. If
you need to obtain a partial result, the fourth parameter \tt{\$skip} and
fifth parameter \tt{\$max} allow you to skip a number of records (the starting
point) and the maximum number to return.


\subsection{QueryForPagedList}
\begin{verbatim}
 public function queryForPagedList($statementName, $parameter=null, $pageSize=10);
\end{verbatim}
We live in an age of information overflow. A database query often returns more
hits than users want to see at once, and our requirements may say that we need
to offer a long list of results a ``page'' at a time. If the query returns
1000 hits, we might need to present the hits to the user in sets of fifty, and
let them move back and forth between the sets. Since this is such a common
requirement, the framework provides a convenience method.

The \tt{TSqlMapPagedList} interface includes methods for navigating through
pages (\tt{nextPage()}, \tt{previousPage()}, \tt{gotoPage(\$pageIndex)}) and
also checking the status of the page (\tt{getIsFirstPage()},
\tt{getIsMiddlePage()}, \tt{getIsLastPage()}, \tt{getIsNextPageAvailable()},
\tt{getIsPreviousPageAvailable()}, \tt{getCurrentPageIndex()},
\tt{getPageSize()}). The total number of records available is not accessible
from the \tt{TSqlMapPagedList} interface, unless a virtual count is defined
using \tt{setVirtualCount(\$value)}, this should be easily accomplished by
simply executing a second statement that counts the expected results.

\begin{mybox}{Tip:}
The \tt{queryForPagedList} method is convenient, but note that a larger set
(up to 3 times the page size) will first be returned by the database provider
and the smaller set extracted by the framework. The higher the page size, the
larger set that will be returned and thrown away. For very large sets, you may
want to use a stored procedure or your own query that uses \tt{\$skip} and
\tt{\$max} as parameters in \tt{queryForList}.
\end{mybox}

\subsection{QueryForMap}
\begin{verbatim}
public function queryForMap($statementName, $parameter=null,
                                    $keyProperty=null, $valueProperty=null);
\end{verbatim}
The \tt{queryForList} methods return the result objects within a \tt{TList} or
array instance. Alternatively, the \tt{queryForMap} returns a TMap or
associative array instance. The value of each entry is one of the result
objects. The key to each entry is indicated by the \tt{\$keyProperty}
parameter. This is the name of the one of the properties of the result object,
the value of which is used as the key for each entry. For example, If you
needed a set of \tt{Employee} objects, you might want them returned as a
\tt{TMap} keyed by each object's \tt{EmployeeNumber} property.

If you don't need the entire result object in your result, you can add the
\tt{\$valueProperty} parameter to indicate which result object property should
be the value of an entry. For example, you might just want the
\tt{EmployeeName} keyed by \tt{EmployeeNumber}.

\subsection{Transaction}
The DataMapper API includes methods to demarcate transactional boundaries. A
transaction can be started, committed and/or rolled back. You can call the
transaction methods from the \tt{TSqlMapper} instance.

\begin{verbatim}
// Begin a transactional session using Adodb transaction API
public function beginTransaction()

// Commit a transaction, uses Adodb transaction API
public function commitTransaction()

// RollBack a transaction, uses Adodb transaction API
public void RollBackTransaction()
\end{verbatim}

\begin{example}\label{example:9.15}
Using transactions
\begin{verbatim}
try
{
    $sqlMap->beginTransaction();
    $item = $sqlMap->queryForObject("getItem", $itemId);
    $item->setDescription($newDescription);
    $sqlMap->update("updateItem", $item);
    $sqlMap->commitTransaction();
}
catch
{
    $sqlMap->rollBackTransaction();
}
\end{verbatim}
\end{example}

\section{Coding Examples}
\begin{example}\label{example:10.1}
Executing Update (insert, update, delete)
\begin{verbatim}
$product = new Product();
$product->setId(1);
$product->setDescription('Shih Tzui');

$key = $sqlMap->insert('insertProduct', $product);
\end{verbatim}
\end{example}

\begin{example}\label{example:10.2}
Executing Query for Object (select)
\begin{verbatim}
$key = 1;
$product = $sqlMap->queryForObject ('getProduct', $key);
\end{verbatim}
\end{example}

\begin{example}\label{example:10.3}
Executing Query for Object (select) With Preallocated Result Object
\begin{verbatim}
$customer = new Customer();

$sqlMap->beginTransaction();

$sqlMap->queryForObject('getCust', $parameter, $customer);
$sqlMap->queryForObject('getAddr', $parameter, $customer);
$sqlMap->commitTransaction();
\end{verbatim}
\end{example}

\begin{example}\label{example:10.4}
Executing Query for List (select)
\begin{verbatim}
$list = $sqlMap->queryForList ('getProductList');
\end{verbatim}
\end{example}

\begin{example}\label{example:10.4}
Executing Query for List (select) With Result Boundaries
\begin{verbatim}
$list = $sqlMap->queryForList ('getProductList', $key, null, 0, 40);
\end{verbatim}
\end{example}

\begin{example}\label{example:10.5}
Executing Query for Paginated List (select)
\begin{verbatim}
$list = $sqlMap->queryForPagedList ('getProductList', null, 10);
$list->nextPage();
$list->previousPage();
\end{verbatim}
\end{example}

\begin{example}\label{example:10.6}
Executing Query for Map
\begin{verbatim}
 $map = $sqlMap->QueryForMap('getProductList', null, 'productCode');
 $product = $map['EST-93'];
\end{verbatim}
\end{example}
