\documentclass{article}
%\usepackage{graphicx}

\usepackage[pdftex]{hyperref}
\usepackage[pdftex]{graphicx}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}


\newtheorem{example}{Example}[section]

\renewcommand{\tt}[1]{\texttt{#1}}

%---------- fonts Type 1 -----------------
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{textcomp}

%------------------------Page set-up-----------------------------------------

\renewcommand{\baselinestretch}{1.25}
\setlength{\hoffset}{-1in}
\setlength{\oddsidemargin}{3.5cm}
\setlength{\evensidemargin}{3.5cm}
\setlength{\topmargin}{0cm}
\setlength{\footskip}{2cm}
\setlength{\headheight}{14pt}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0cm}
\setlength{\parskip}{0.75\baselineskip}

\hypersetup{colorlinks, linkcolor=blue, pdfstartview={FitH}}

%------------------------------------------------------------------------------

\newsavebox{\fmboxb}
\newenvironment{mybox}[1]
    {\vspace{-2mm}\begin{center}\begin{lrbox}{\fmboxb}\hspace{2mm}
        \begin{minipage}{0.85\textwidth} \vspace{2mm}\small \textbf{#1}}
    {  \vspace{2mm} \end{minipage}
    \hspace{2mm}\end{lrbox}\fbox{\usebox{\fmboxb}}\end{center}}

%----------------- TITLE --------------

\title{\vspace{-2.5cm} \bfseries SQLMap PHP DataMapper Tutorial
    \thanks{Copyright 2006. All Rights Reserved.}
}
\author{Wei Zhuo}
\date{\today}


%-------------- BEGIN DOCUMENT ------------------


\begin{document}

\maketitle

This tutorial takes an ``over-the-shoulder'' Cookbook approach. We'll define a
simple data access problem and use SQLMap to solve it for us.


\section*{Legal Notice}

Copies of this document may be made for your own use and for distribution to
others, provided that you do not charge any fee for such copies and further
provided that each copy contains this Copyright Notice, whether distributed in
print or electronically.

This document is largely based on the iBATIS.NET -- DataMapper Application
Tutorial.

\subsection*{License Information}
SQLMap for PHP is free software released under the terms of the following BSD
license. Copyright 2004-2006, PradoSoft (http://www.pradosoft.com) All rights
reserved.

\subsection*{Disclaimer}
SQLMap PHP DataMapper MAKES NO WARRANTIES, EXPRESS OR IMPLIED, AS TO THE
INFORMATION IN THIS DOCUMENT. The names of actual companies and products
mentioned herein may be the trademarks of their respective owners.

\subsection*{Remark}
Original writing by Clinton Begin, Ted Husted and Gilles Bayon.

\include{tut1}
\include{tut2}
\include{tut3}


\end{document}
